# DM termpaper accuracy U.S. election polls

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages needed
p_needed <-
  c("viridis", "stargazer", "foreign", "readxl", "readr", "dplyr", "strptime", "tidyr", "ggplot2", "ggthemes", "gridExtra")

# packages installed on your computer
packages <- rownames(installed.packages())

p_to_install <- p_needed[!(p_needed %in% packages)]

# install if not on computer
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}

sapply(p_needed, require, character.only = TRUE)

# This is an option for stargazer tables
stargazer_opt <- ifelse(knitr::is_latex_output(), "latex", "html")
```

# Load datasets
```{r load datasets, warning = FALSE, message = FALSE}
# 538 polls 2020 (https://projects.fivethirtyeight.com/2020-election-forecast/)
polls2020 <- read_csv("raw_data/president_polls_2020.csv") 
#head(polls2020)

# 538 polls 2016 (https://projects.fivethirtyeight.com/2016-election-forecast/) 
polls2016 <- read_csv("raw_data/president_general_polls_2016.csv") 
#head(polls2016)

# 538 pollster info (https://github.com/fivethirtyeight/data/tree/master/pollster-ratings) including additional coded information
pollster2020 <- read_excel("raw_data/2020_pollster-stats-full_coded.xlsx")
#head(pollster2020)

pollster2016 <- read_excel("raw_data/2016_pollster-stats-full_coded.xlsx")
#head(pollster2016)

# election results (2016 + 2020 - state; https://github.com/MEDSL/election_night2020/tree/main/results)
results_16_20 <- read_excel("raw_data/timestamp_results_nyt_call.xlsx")
#head(results_16_20)

# covid data (https://github.com/MEDSL/election_night2020/tree/main/context/covid/raw)
covid_state <- read_rds("raw_data/covid_state_ts.rds")
#head(covid)

```

# Prepare and merge datasets
## Election results dataset
```{r election election results}

# cleaning and filtering
results_16_20$timestamp <- as.numeric(results_16_20$timestamp) # transform to character

max_results <- results_16_20 %>%
                  group_by(state) %>%
                  filter(timestamp == max(timestamp, na.rm = TRUE)) # filter to newest (max) timestamp

# head(max_results)

results_cleaned <- max_results[c("votes", "trumpd", "bidenj", "race_rating", "state_name", "state_id", "electoral_votes", "absentee_votes", "reporting_display", "reporting_value", "leader_margin_name_display", "leader_party_id", "votes2016", "margin2016", "clinton2016", "trump2016")] # keep only these variables

# add national level results, taken from: https://uselectionatlas.org/RESULTS/ 
result <- rbind(results_cleaned, c(155509279, 0.468, 0.512, "NA", "U.S.", "U.S.", 538, "NA", "NA", 0, "Biden + 4", "democrat", 128838830, -2.1, 65853677, 62985153)) # 2020

# add percentages for 2016 election (Clinton + Trump)
table(result$state_id)
result$trump2016p <- c(0.512, 0.62, 0.605, 0.48, 0.314, 0.432, 0.409, 0.04, 0.417, 0.486, 0.503, 0.3, 0.511, 0.592, 0.383, 0.564, 0.56, 0.625, 0.58, 0.328, 0.339, 0.448, 0.472, 0.449, 0.563, 0.578, 0.556, 0.498, 0.629, 0.587, 0.464, 0.41, 0.4, 0.455, 0.365, 0.513, 0.653, 0.39, 0.481, 0.389, 0.549, 0.615, 0.607, 0.52, 0.45, 0.444, 0.302, 0.368, 0.472, 0.678, 0.681, 0.459) 

result$clinton2016p <- c(0.365, 0.343, 0.336, 0.445, 0.614, 0.481, 0.545, 0.908, 0.53, 0.474, 0.452, 0.622, 0.417, 0.274, 0.552, 0.374, 0.356, 0.326, 0.384, 0.6, 0.603, 0.478, 0.47, 0.464, 0.378, 0.4, 0.354, 0.461, 0.272, 0.337, 0.468, 0.549, 0.482, 0.479, 0.59, 0.432, 0.289, 0.5, 0.474, 0.544, 0.406, 0.317, 0.347, 0.431, 0.271, 0.497, 0.566, 0.525, 0.464, 0.261, 0.218, 0.48)

# head(result)

result$winner2020_biden <- ifelse(result$bidenj>result$trumpd, 
                            1, 
                            ifelse(is.na(result$bidenj), NA, 0)) # winner is Biden

result$winner2016_clinton <- ifelse(result$clinton2016p>result$trump2016p, 
                            1, 
                            ifelse(is.na(result$clinton2016p), NA, 0)) # winner is Clinton

result[52, 20] <- 0 # fix the popular vote for 2016 ^^ 
  
# table(result$winner2016_trump, result$state_name) 
# table(result$winner2020_trump, result$state_name) 

# add electoral votes (states + national), taken from: https://uselectionatlas.org/RESULTS/ 
result$electoral_votes2020_b <- c(0, 0, 6, 11, 55, 9, 7, 3, 3, 0, 16, 4, 0, 0, 20, 0, 0, 0, 0, 11, 10, 3, 16, 10, 0, 0, 0, 0, 0, 1, 4, 14, 5, 6, 29, 0, 0, 7, 20, 4, 0, 0, 0, 0, 0, 13, 3, 12, 10, 0, 0, 306)
result$electoral_votes2020_t <- c(3, 9, 0, 0, 0, 0, 0, 0, 3, 29, 0, 0, 6, 4, 0, 11, 6, 8, 8, 0, 0, 1, 0, 0, 10, 6, 3, 15, 3, 4, 0, 0, 0, 0, 0, 18, 7, 0, 0, 0, 9, 3, 11, 38, 6, 0, 0, 0, 0, 5, 3, 232)

result$electoral_votes2016_c <- c(0, 0, 0, 0, 55, 9, 7, 3, 3, 0, 0, 3, 0, 0, 20, 0, 0, 0, 0, 11, 10, 3, 0, 10, 0, 0, 0, 0 , 0, 0, 4, 14, 5, 6, 29, 0, 0, 7, 0, 4, 0, 0, 0, 0, 0, 13, 3, 8, 0, 0, 0, 227)
result$electoral_votes2016_t <- c(3, 9, 6, 11, 0, 0, 0, 0, 0, 29, 16, 0, 6, 4, 0, 11, 6, 8, 8, 0, 0, 1, 16, 0, 10, 6, 3, 15, 3, 5, 0, 0, 0, 0, 0, 18, 7, 0, 20, 0, 9, 3, 11, 38, 6, 0, 0, 0, 0, 5, 3, 304)

# head(result)

```

## 2020 polls
```{r 2020 polls}

trump <- polls2020[polls2020$candidate_id==13254,] # subset to Trump
trump2 <- trump[-c(3, 6:33)]

biden <- polls2020[polls2020$candidate_id==13256,] # subset to Biden

polls_2020_combined_test <- trump2 %>% # generate new dataset
  left_join(biden, by = c("poll_id", "state", "pollster_id", "question_id")) # merge subsets (Trump to Biden)

polls_2020_combined <- polls_2020_combined_test[-c(1)]

# head(polls_2020_combined)
polls_2020_combined2 <- polls_2020_combined[c("poll_id", "state", "pollster_id", "pollster", "sponsors", "pollster_rating_id", "fte_grade", "sample_size", "population", "population_full", "methodology", "start_date", "end_date", "nationwide_batch", "answer.x", "candidate_id.x", "candidate_party.x", "pct.x", "answer.y", "candidate_id.y", "candidate_party.y", "pct.y")] # keep these variables

# head(polls_2020_combined2)
```

## 2020 pollsters 
```{r 2020 pollsters}

# head(pollster2020)

pollster_2020_cleaned <- pollster2020[c("Pollster", "Pollster Rating ID", "Nr of Polls", "NCPP / AAPOR / Roper",	
"Live Caller With Cellphones", "Banned by 538", "Races Called Correctly", "Herding Penalty", "English", "Spanish", "Weight by age", "Weight by gender", "Weight by education", "Weight by race", "Weight by partisanship/party registration", "Weight by region/district", "Weight by income", "Weight by other")] # keep these variables
# head(pollster_2020_cleaned)

```

## merge 2020 polls and pollsters
```{r merge 2020 polls and pollsters}

colnames(pollster_2020_cleaned)[colnames(pollster_2020_cleaned)=="Pollster Rating ID"]<-"pollster_rating_id" # make sure that variable names for merging match
colnames(pollster_2020_cleaned)[colnames(pollster_2020_cleaned)=="Pollster"]<-"pollster"

pp2020 <- polls_2020_combined2 %>% # generate new, merged dataset 
  left_join(pollster_2020_cleaned, by = c("pollster_rating_id", "pollster"))

# head(pp2020) 
# nrow(pp2020) # 6020

pp2020 <- pp2020[!is.na(pp2020$pct.x), ] # delete observations for which Biden is not polled
pp2020 <- pp2020[!is.na(pp2020$pct.y), ] # delete observations for which Trump is not polled

```

## 2016 polls
```{r polls 2016}
p2016 <- polls2016[polls2016$type=="polls-only",] # polls without adjustment for other variables

# head(p2016)

polls_2016_cleaned <- p2016[c("state", "startdate", "enddate", "pollster", "grade", "samplesize", "population", "rawpoll_clinton", "rawpoll_trump", "poll_id")] # keep these variables

# head(polls_2016_cleaned)
```

## 2016 pollsters
```{r pollsters 2016}
# head(pollster2016) 

pollster_2016_cleaned <- pollster2016[c("Pollster", "Nr of Polls", "NCPP / AAPOR / Roper",	"Live Caller With Cellphones", "Internet","Banned by 538", "Races Called Correctly", "English", "Spanish", "Weight by age", "Weight by gender", "Weight by education", "Weight by race", "Weight by partisanship/party registration", "Weight by region/district", "Weight by income", "Weight by other")] # keep these variables

# head(pollster_2016_cleaned)
```

## merge 2016 polls and pollsters
```{r merge 2016 polls and pollsters}
colnames(pollster_2016_cleaned)[colnames(pollster_2016_cleaned)=="Pollster"]<-"pollster" # make sure that variable names for merging match

pp2016 <- pollster_2016_cleaned %>% # generate new, merged dataset
  left_join(polls_2016_cleaned, by = c("pollster"))

# head(pp2016)
# nrow(pp2016) # 4441 cases

pp2016 <- pp2016[!is.na(pp2016$rawpoll_clinton), ] # delete observations for which Clinton is not polled
pp2016 <- pp2016[!is.na(pp2016$rawpoll_trump), ] # delete observations for which Trump is not polled
# nrow(pp2016) # 4208 cases
```

## merge election results with 2016 dataset
```{r merge election results with 2016 dataset}
# head(pp2016)
# head(result)
final_2016 <- pp2016 %>% # generate new, merged dataset
  left_join(result, by = c("state"="state_name"))

# head(final_2016)

table(final_2016$clinton2016p, useNA = "ifany") # there are missings -> Maine and Nebraska has district polls but we have only state level results!
table(result$state_name)
table(pp2016$state)

# nrow(final_2016) # 4208
final_2016 <- final_2016[!is.na(final_2016$state_id), ] # Maine and Nebraska districts omitted
# nrow(final_2016) # 4179 

```

## merge election results with 2020 dataset
```{r merge election results with 2020 dataset}
# head(pp2020)
# head(result)

final_2020 <- pp2020 %>% # generate new, merged dataset
  left_join(result, by = c("state"="state_name"))

# head(final_2020)

table(final_2020$bidenj, useNA = "ifany") # there are missings -> Maine and Nebraska has district polls but we have only state level results!
table(result$state_name)
table(pp2020$state)
# nrow(final_2020) # 6020
final_2020 <- final_2020[!is.na(final_2020$state_id), ] # Maine and Nebraska districts omitted
# nrow(final_2020) # 5700

# delete 2016 cases in final_2020 and delete 2020 cases in final_2016?
final_2020$trumpd <- as.numeric(final_2020$trumpd)
final_2020$bidenj <- as.numeric(final_2020$bidenj)
summary(final_2020$trumpd)
summary(final_2020$bidenj) 
```

## recode variables (methodology, ...)
```{r recode methodology}

#table(final_2020$methodology) 

# recoding the methodology to phone, online, mix, other (pure categories)
final_2020$meth_cat <- as.numeric(as.factor(final_2020$methodology)) #final_2020$methodology 

final_2020$meth_cat[final_2020$meth_cat == 1] <- "phone"
final_2020$meth_cat[final_2020$meth_cat == 2] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 3] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 4] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 5] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 6] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 7] <- "phone"
final_2020$meth_cat[final_2020$meth_cat == 8] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 9] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 10] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 11] <- "other"
final_2020$meth_cat[final_2020$meth_cat == 12] <- "online"
final_2020$meth_cat[final_2020$meth_cat == 13] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 14] <- "mix"
final_2020$meth_cat[final_2020$meth_cat == 15] <- "other"


table(final_2020$meth_cat)

```
# coding grade and live caller for 2020
```{r grade variable}
# 2016 grade 
final_2016$grade_new <- final_2016$grade

final_2016$grade_new[final_2016$grade_new == "A-"] <- "A"
final_2016$grade_new[final_2016$grade_new == "A+"] <- "A"
final_2016$grade_new[final_2016$grade_new == "B-"] <- "B"
final_2016$grade_new[final_2016$grade_new == "B+"] <- "B"
final_2016$grade_new[final_2016$grade_new == "C-"] <- "C"
final_2016$grade_new[final_2016$grade_new == "C+"] <- "C"

# table(final_2016$grade_new, useNA = "ifany")

# 2020 grade 
final_2020$fte_grade_new <- final_2020$fte_grade
final_2020$fte_grade_new[final_2020$fte_grade_new == "A+"] <- "A"
final_2020$fte_grade_new[final_2020$fte_grade_new == "A-"] <- "A"
final_2020$fte_grade_new[final_2020$fte_grade_new == "A/B"] <- "B" # more conservative coding
final_2020$fte_grade_new[final_2020$fte_grade_new == "B-"] <- "B"
final_2020$fte_grade_new[final_2020$fte_grade_new == "B+"] <- "B"
final_2020$fte_grade_new[final_2020$fte_grade_new == "B/C"] <- "C"
final_2020$fte_grade_new[final_2020$fte_grade_new == "C+"] <- "C"
final_2020$fte_grade_new[final_2020$fte_grade_new == "C-"] <- "C"
final_2020$fte_grade_new[final_2020$fte_grade_new == "C/D"] <- "D"
final_2020$fte_grade_new[final_2020$fte_grade_new == "D-"] <- "D"

# table(final_2020$fte_grade_new, useNA = "ifany")

# alternative recoding 2020 grade
final_2020$fte_grade_newalt <- final_2020$fte_grade
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "A+"] <- "A"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "A-"] <- "A"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "A/B"] <- "A" # more conservative coding
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "B-"] <- "B"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "B+"] <- "B"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "B/C"] <- "B"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "C+"] <- "C"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "C-"] <- "C"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "C/D"] <- "C"
final_2020$fte_grade_newalt[final_2020$fte_grade_newalt == "D-"] <- "D"

# table(final_2020$fte_grade_newalt, useNA = "ifany")

# live caller
final_2020$cellphones <- final_2020$`Live Caller With Cellphones`
final_2020$cellphones[final_2020$cellphones == "yes*"] <- "yes"
final_2020$cellphones[final_2020$cellphones == "partial"] <- "yes"

# table(final_2020$cellphones, useNA = "ifany")

``` 


## prepare winner for polls
```{r winner-variable for polls}

# 2016
final_2016$clinton_new <- final_2016$clinton2016p * 100 # multiply to match rawpoll numbers
final_2016$trump_new <- final_2016$trump2016p * 100


final_2016$winner2016poll_clinton <- ifelse(final_2016$rawpoll_clinton>final_2016$rawpoll_trump, 
                            1, 
                            ifelse(is.na(final_2016$rawpoll_clinton), NA, 0)) # 1 if Clinton is winner 

# table(final_2016$winner2016poll_clinton, useNA = "ifany")

# 2020

final_2020$trump_new <- final_2020$trumpd * 100 # multiply to match rawpoll numbers
final_2020$biden_new <- final_2020$bidenj * 100

final_2020$winner2020poll_biden <- ifelse(final_2020$pct.y>final_2020$pct.x, 
                            1, 
                            ifelse(is.na(final_2020$pct.y), NA, 0)) # 1 if Biden is winner (x is Trump, y is Biden)

# table(final_2020$winner2020poll_biden, useNA = "ifany") 
```

# Code accuracy variables
## accuracy as right winner
```{r accuracy - winner}

# 2016
# table(final_2016$winner2016_clinton, useNA = "ifany")
#table(final_2016$winner2016poll_clinton, useNA = "ifany")

final_2016$accuracy_poll <- ifelse(final_2016$winner2016_clinton == final_2016$winner2016poll_clinton, 
                            1, 
                            ifelse(is.na(final_2016$winner2016_clinton), NA, 0)) # 1 if poll predicts correctly (winner = winner, loser = loser) 

table(final_2016$accuracy_poll, useNA = "ifany")
# most polls predicted winner/loser correctly

# 2020
# table(final_2020$winner2020_biden, useNA = "ifany")
# table(final_2020$winner2020poll_biden, useNA = "ifany")

final_2020$accuracy_poll <- ifelse(final_2020$winner2020_biden == final_2020$winner2020poll_biden, 
                            1, 
                            ifelse(is.na(final_2020$winner2020_biden), NA, 0)) # 1 if poll predicts correctly (winner = winner, loser = loser) 

# table(final_2020$accuracy_poll, useNA = "ifany")
# more polls than 2016 predict winner/loser correctly 
# table(final_2020$accuracy_poll, final_2020$state_id, useNA = "ifany")
# incorrect predictions mostly in AZ, FL, GA, IA, MI, NC, OH, TX -> swing states

```

## accuracy AAPOR 
```{r accuracy AAPOR}
# 2016
final_2016$margin_poll2016 <- final_2016$rawpoll_clinton - final_2016$rawpoll_trump
# extreme values of -46 and + 86 are legitimate

final_2016$margin_result2016 <- final_2016$clinton_new - final_2016$trump_new

# table(final_2016$margin_poll2016, useNA = "ifany")
# table(final_2016$margin_result2016, useNA = "ifany")

final_2016$accuracy_aapor <- final_2016$margin_poll2016 - final_2016$margin_result2016

# table(final_2016$accuracy_aapor, useNA = "ifany")
range(final_2016$accuracy_aapor)

# 2020
final_2020$margin_poll2020 <- final_2020$pct.y - final_2020$pct.x  # Biden (y) - Trump (x)

final_2020$margin_result2020 <- final_2020$biden_new - final_2020$trump_new

final_2020$accuracy_aapor <- final_2020$margin_poll2020 - final_2020$margin_result2020

# table(final_2020$accuracy_aapor, useNA = "ifany")
range(final_2020$accuracy_aapor) 

```

## accuracy AAPOR absolute
```{r accuracy AAPOR absolute}
# 2016
final_2016$accuracy_aapor_abs <- abs(final_2016$margin_poll2016 - final_2016$margin_result2016)

range(final_2016$accuracy_aapor_abs)

# 2020
final_2020$accuracy_aapor_abs <- abs(final_2020$margin_poll2020 - final_2020$margin_result2020)

range(final_2020$accuracy_aapor_abs)
```

## format time
```{r format time}
# 2016
#table(final_2016$startdate)
class(final_2016$startdate)
final_2016$startdate2 <- as.Date(as.character(final_2016$startdate), "%m/%d/%Y")
#table(final_2016$startdate2)
class(final_2016$startdate2)

final_2016 <- final_2016[final_2016$startdate2 > as.Date("2015-12-31"),] # subset data to election year

# 2020
#table(final_2020$start_date)
final_2020$start_date2 <- as.Date(as.character(final_2020$start_date), "%m/%d/%y")
class(final_2020$start_date2)
#table(final_2020$start_date2, useNA = "ifany")
final_2020$start_date3 <- strftime(final_2020$start_date2, format = "%V")
#table(final_2020$start_date3, useNA = "ifany")
final_2020$start_date3 <- as.numeric(final_2020$start_date3)

final_2020 <- final_2020[final_2020$start_date2 > as.Date("2019-12-31"),] # subset data to election year

```

## merge Covid data
```{r Covid}
covid_state <- covid_state[c(2,4:11)] # delete unnecessary variables

final_2020c <- final_2020 %>% # generate new, merged dataset
  left_join(covid_state, by = c("start_date3"="week", "state_id" ="state_po"))

#table(final_2020c$state)
#range(final_2020c$cases_per100k, na.rm = T) # 0 -730
```

```{r save dataset}
save(final_2016, file = "merged_data/final_2016.RData") 

save(final_2020c, file = "merged_data/final_2020.RData")

```

# Descriptive statistics
## Get overview of dataset
### 2016

```{r descriptive statistics 2016}
nrow(final_2016) # number of observations: 4124

length(unique(final_2016$pollster)) # 195 pollsters

table(final_2016$grade_new, useNA = "ifany") # most pollster A and B, even with conservative coding

table(final_2016$Internet, useNA = "ifany") # majority includes Internet
table(final_2016$`Live Caller With Cellphones`, useNA = "ifany") # minority uses cellphone

table(final_2016$population) # most include likely voters 

table(final_2016$accuracy_poll) # more than half of polls predict correctly (winner and loser)

summary(final_2016$accuracy_aapor) # mean is 4, range -80 to 42
summary(final_2016$accuracy_aapor_abs) # mean is 6.7, range 0 to 80 (by definition always positive)

summary(final_2016$samplesize) 

table(final_2016$`Weight by education`, final_2016$grade_new, useNA = "ifany")

table(final_2016$`Weight by education`, final_2016$state_id, useNA = "ifany")


aggregate(x = final_2016$samplesize,                
          by = list(final_2016$grade_new),              
          FUN = mean) 
```

### 2020

```{r descriptive statistics 2020}
nrow(final_2020c) # number of observations: 5700

length(unique(final_2020c$pollster)) # 212 pollsters

table(final_2020c$fte_grade_new, useNA = "ifany") # majority of pollsters received D rating

table(final_2020c$meth_cat, useNA = "ifany") # majority includes Internet (and some phone)

table(final_2020c$cellphones, useNA = "ifany") # minority uses cellphone

table(final_2020c$population) # most include likely voters 

table(final_2020c$accuracy_poll, useNA = "ifany") # a majority of polls predict correctly (winner and loser)

table(final_2020c$accuracy_poll, final_2020c$state, useNA = "ifany")

summary(final_2020c$accuracy_aapor) # mean is 5, range -21 to 25
summary(final_2020c$accuracy_aapor_abs) # mean is 5.6, range 0 to 25 (by definition always positive)

# sample size
summary(final_2020c$sample_size) 

table(final_2020c$`Weight by education`, final_2020c$fte_grade_new, useNA = "ifany")

table(final_2020c$`Weight by education`, final_2020c$state_id, useNA = "ifany")

plot(
  density(final_2020c$accuracy_aapor, na.rm=T))
plot(
  density(final_2020c$accuracy_aapor_abs, na.rm=T))

table(final_2016$state_id)

summary(final_2020c$accuracy_aapor_abs)

table(final_2020c$meth_cat, final_2020c$fte_grade_new, useNA = "ifany")
table(final_2020c$`Weight by education`, final_2020c$fte_grade_new, useNA = "ifany")

aggregate(x = final_2020c$sample_size,                
          by = list(final_2020c$fte_grade_new),              
          FUN = mean) # largest average sample size for D-rated pollsters
```

# Plots
## prepare plots 
```{r prepare plots accuracy}
# 2016

swing2016 <- subset(final_2016, state_id == "AZ" | state_id == "CO" | state_id == "FL" | state_id == "GA" | state_id == "IA" | state_id == "ME" | state_id == "MI" | state_id == "MN" | state_id == "NC" | state_id == "NH" | state_id == "NV" | state_id == "OH" | state_id == "PA" | state_id == "VA" | state_id == "TX" | state_id == "WI") # based on 'toss-up' and 'leaning' rating of NYT election results and FTE

red2016 <- subset(final_2016, state_id == "AL" | state_id == "AR" | state_id == "ID" | state_id == "KY" | state_id == "LA" | state_id == "MS" | state_id == "NE" | state_id == "ND" | state_id == "OK" | state_id == "SD" | state_id == "TN" | state_id == "WV" | state_id == "WY") # based on solid-rep rating of NYT election results


blue2016 <- subset(final_2016, state_id == "CA" | state_id == "HI" | state_id == "DE" | state_id =="DC" | state_id == "CT" | state_id =="MD" | state_id =="MA" | state_id =="IL" | state_id =="RI" | state_id == "OR" | state_id =="NY" | state_id =="NM" | state_id == "NJ" | state_id == "WA" | state_id =="VT") # based on solid-dem rating of NYT election results

national2016 <- subset(final_2016, state_id == "U.S.")


# 2020
swing2020 <- subset(final_2020c, state_id == "AZ" | state_id == "CO" | state_id == "FL" | state_id == "GA" | state_id == "IA" | state_id == "ME" | state_id == "MI" | state_id == "MN" | state_id == "NC" | state_id == "NH" | state_id == "NV" | state_id == "OH" | state_id == "PA" | state_id == "VA" | state_id == "TX" | state_id == "WI") # based on 'toss-up' and 'leaning' rating of NYT election results and FTE

red2020 <-  subset(final_2020c, state_id == "AL" | state_id == "AR" | state_id == "ID" | state_id == "KY" | state_id == "LA" | state_id == "MS" | state_id == "NE" | state_id == "ND" | state_id == "OK" | state_id == "SD" | state_id == "TN" | state_id == "WV" | state_id == "WY") # based on solid-rep rating of NYT election results


blue2020 <- subset(final_2020c, state_id == "CA" | state_id == "HI" | state_id == "DE" | state_id =="DC" | state_id == "CT" | state_id =="MD" | state_id =="MA" | state_id =="IL" | state_id =="RI" | state_id == "OR" | state_id =="NY" | state_id =="NM" | state_id == "NJ" | state_id == "WA" | state_id =="VT") # based on solid-dem rating of NYT election results

national2020 <- subset(final_2020c, state_id == "U.S.")

```

## plots accuracy
```{r plots accuracy, warning = FALSE, message = FALSE}
# barplot mean accuracy
bar_accuracy <- cbind(c(mean(final_2016$accuracy_aapor_abs), mean(national2016$accuracy_aapor_abs), mean(red2016$accuracy_aapor_abs), mean(blue2016$accuracy_aapor_abs), mean(swing2016$accuracy_aapor_abs)),
            c(mean(final_2020c$accuracy_aapor_abs), mean(national2020$accuracy_aapor_abs), mean(red2020$accuracy_aapor_abs), mean(blue2020$accuracy_aapor_abs), mean(swing2020$accuracy_aapor_abs))) # take means for dataset and subsets

png(file="tables_figures_and_graphs/barplot_accuracy.png", width=600, height=350) 
x <- barplot(bar_accuracy, 
  beside=T,                                      
  names.arg=c('2016 U.S. presidential election','2020 U.S. presidential election'),  # Säulengruppen beschriften
  col=c('darkorange','darkseagreen','brown2','cornflowerblue', 'mediumpurple'),  
  ylim=c(0, 16),                               
  xpd=FALSE,                                   
  main = "Poll accuracy by election",
  xlab = "",
  ylab = "Absolute error on poll margin"
)
par(cex = 1) # sets font size of numbers
y<-as.matrix(bar_accuracy)
text(x, y+2, labels=as.character(round(y,2)))
legend("topright",        # Add legend to plot
       legend = c('Overall','National','Republican','Democratic', 'Swing'),
       col = c('darkorange','darkseagreen','brown2','cornflowerblue', 'mediumpurple'),
       pch = 16,
       cex = 1)
dev.off()

# swing states
s2016 <- swing2016 %>% group_by(state_id) %>% summarise(means = mean(accuracy_aapor_abs)) %>% ggplot(aes(x = state_id, y = means, fill = state_id)) +
    geom_bar(stat = "identity", fill = "#999999", color = "black") +
    theme_classic() +
    labs(
        x = "",
        y = "Absolute error on poll margin",
       stat="identity",
        title = paste(
            "Absolute poll accuracy in swing states in 2016"
        )) +
    geom_text(aes(label=round(means, 2)), position=position_dodge(width=0.9), vjust=-0.25) +
  scale_y_continuous(limits = c(0, 9)) 
   # png('tables_figures_and_graphs/accuracy_swing2016.png',width=600,height=350,units="px",bg = "transparent")


s2020 <- swing2020 %>% group_by(state_id) %>% summarise(means = mean(accuracy_aapor_abs)) %>% ggplot(aes(x = state_id, y = means, fill = state_id)) +
    geom_bar(stat = "identity", fill = "#999999", color = "black") +
    theme_classic() +
    labs(
        x = "",
        y = "Absolute error on poll margin",
        title = paste(
            "Absolute poll accuracy in swing states in 2020"
        ) # + ylim(0, 15)
    ) +
    geom_text(aes(label=round(means, 2)), position=position_dodge(width=0.9), vjust=-0.25) +
  scale_y_continuous(limits = c(0, 9)) 
 # png('tables_figures_and_graphs/accuracy_swing2020.png',width=600,height=350,units="px",bg = "transparent")

png(file="tables_figures_and_graphs/swingstates.png", width=900, height=400) 
grid.arrange(s2016, s2020, nrow = 1, ncol = 2)
dev.off()

```
## plots accuracy over time
```{r plot time accuracy}
# 2016
na16 <- lm(accuracy_aapor_abs ~ grade_new + startdate2, data = final_2016) # remove NAs

ggplot(na16, aes(y = accuracy_aapor_abs, x = startdate2)) +
  geom_point(aes(colour = factor(grade_new)), alpha = 0.3) +
  facet_wrap("grade_new", nrow = 3) +
   geom_smooth(span = 0.5, colour = "black") +
   scale_y_continuous("Absolute error on poll margin") +
   ggtitle("Poll accuracy and pollster grade 2016") +
   labs(x = "Start date of survey", color = "Pollster grade") + 
   png('tables_figures_and_graphs/accuracy2016.png',width=600,height=450,units="px",bg = "transparent")

# 2020
na20 <- lm(accuracy_aapor_abs ~ fte_grade_new + start_date2, data = final_2020c) # remove NAs

ggplot(na20, aes(y = accuracy_aapor_abs, x = start_date2)) +
  geom_point(aes(colour = factor(fte_grade_new)), alpha = 0.3) +
  facet_wrap("fte_grade_new", nrow = 3) +
   geom_smooth(span = 0.5, colour = "black") +
   scale_y_continuous("Absolute error on poll margin") +
   ggtitle("Poll accuracy and pollster grade 2020") +
   labs(x = "Start date of survey", color = "Pollster grade") + 
   png('tables_figures_and_graphs/accuracy2020.png',width=600,height=450,units="px",bg = "transparent")

# Plot with alternative recoding of FTE grade (not conservative)
na20alt <- lm(accuracy_aapor_abs ~ fte_grade_newalt + start_date2, data = final_2020c) # remove NAs

ggplot(na20alt, aes(y = accuracy_aapor_abs, x = start_date2)) +
  geom_point(aes(colour = factor(fte_grade_newalt)), alpha = 0.3) +
  facet_wrap("fte_grade_newalt", nrow = 3) +
   geom_smooth(span = 0.5, colour = "black") +
   scale_y_continuous("Absolute error on poll margin") +
   ggtitle("Poll accuracy and pollster grade 2020 (alternative grade)") +
   labs(x = "Start date of survey", color = "Pollster grade") + 
   png('tables_figures_and_graphs/accuracy2020_altgrade.png',width=600,height=450, units="px",bg = "transparent")

```

# Regression 2016

```{r regression 2016, warning = FALSE, message = FALSE}
# Model 1
m1 <- lm(accuracy_aapor_abs ~ `Live Caller With Cellphones` + Internet, data = final_2016) 
#summary(m1)

# Model 2
m2 <- lm(accuracy_aapor_abs ~ `Live Caller With Cellphones` + Internet + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age`, data = final_2016) 
#summary(m2)

# Model 3
m3 <- lm(accuracy_aapor_abs ~ `Live Caller With Cellphones` + Internet + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ grade_new + population, data = final_2016) 
#summary(m3)

# subset for time
sub2016 <- final_2016[final_2016$startdate2 > as.Date("2016-09-30"),] # subset data to October and November

# Model 4 (Subset time)
m4 <- lm(accuracy_aapor_abs ~ `Live Caller With Cellphones` + Internet + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ grade_new + population, data = sub2016)
#summary(m4) 

# regression table
stargazer(m1, m2, m3, m4, 
          type = "html",
          covariate.labels = c("Live Caller: Yes", "Internet: Yes", "Weight By Education", "Weight By Gender", "Weight By Race", "Weight By Age", "Weight By Income", "Weight By Region/District", "Weight By Partisanship", "Grade B", "Grade C", "Grade D", "Likely Voters", "Registered Voters"), 
          header = F,
          label = "tab:regression1", 
          title = "2016 Election: Linear Regression",
          keep.stat = c("n", "rsq", "adj.rsq"), 
          no.space = T, # no space between lines
          dep.var.caption = "", 
          dep.var.labels = "Dependent variable: Absolute error on the poll margin",
          out="tables_figures_and_graphs/reg2016.htm")

```

# Regression 2020

```{r regression 2020, warning = FALSE, message = FALSE}
# Model 5
m5 <- lm(accuracy_aapor_abs ~ cellphones + meth_cat, data = final_2020c) # ref is mixed
#summary(m5)

# Model 6
m6 <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age`, data = final_2020c) 
#summary(m6)

# Model 7
m7 <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ final_2020c$`Weight by other` + fte_grade_new + population, data = final_2020c)
# population a is ref category
#summary(m7)

# subset for time
sub2020 <- final_2020c[final_2020c$start_date2 > as.Date("2020-09-30"),] # subset data to October and November

# Model 8 (Subset time)
m8 <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ fte_grade_new + population, data = sub2020)
#summary(m8) 

# Model 9
m9 <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ final_2020c$`Weight by other` + fte_grade_new + population + cases_per100k, data = final_2020c) # covid model
#summary(m9)

# regression table
stargazer(m5, m6, m7, m8, m9, 
          type = "html",
          covariate.labels = c("Cellphones: Yes", "Online", "Other", "Phone", "Weight by Education", "Weight by Gender", "Weight by Race", "Weight by Age", "Weight by Income", "Weight by Region/District", "Weight by Partisanship", "Weight by Other","Grade B", "Grade C", "Grade D", "Likely Voters", "Registered Voters", "Voters", "Covid cases/100k"), 
          header = F, # To surpress the stargazer comment
          label = "tab:regression2", # If you produce pdf output you can even cross-reference your table in your text.
          title = "2020 Election: Linear Regression",
          keep.stat = c("n", "rsq", "adj.rsq"), 
          no.space = T, # no space between lines
          dep.var.caption = "", 
          dep.var.labels = "Dependent variable: Absolute error on the poll margin",
          out="tables_figures_and_graphs/reg2020.htm")

```

```{r regression 2020 alt, warning = FALSE, message = FALSE}
# Model 7 alt
m7alt <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ final_2020c$`Weight by other` + fte_grade_newalt + population, data = final_2020c)
# population a is ref category
#summary(m7alt)

# Model 8 alt (Subset time)
m8alt <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ fte_grade_newalt + population, data = sub2020)
#summary(m8alt) 

# Model 9 alt
m9alt <- lm(accuracy_aapor_abs ~ cellphones + meth_cat + `Weight by education` + `Weight by gender` + `Weight by race` + `Weight by age` + `Weight by income` + `Weight by region/district` + `Weight by partisanship/party registration`+ final_2020c$`Weight by other` + fte_grade_newalt + population + cases_per100k, data = final_2020c) # covid model
#summary(m9alt)

# regression table
stargazer(m5, m6, m7alt, m8alt, m9alt,
          type = "html",
          covariate.labels = c("Cellphones: Yes", "Online", "Other", "Phone", "Weight by Education", "Weight by Gender", "Weight by Race", "Weight by Age", "Weight by Income", "Weight by Region/District", "Weight by Partisanship", "Weight by Other","Grade B", "Grade C", "Grade D", "Likely Voters", "Registered Voters", "Voters", "Covid cases/100k"),
          header = F, # To surpress the stargazer comment
          label = "tab:regression3", # If you produce pdf output you can even cross-reference your table in your text.
          title = "2020 Election: Linear Regression (alternative grade)",
          keep.stat = c("n", "rsq", "adj.rsq"), 
          no.space = T, # no space between lines
          dep.var.caption = "", 
          dep.var.labels = "Dependent variable: Absolute error on the poll margin",
          out="tables_figures_and_graphs/reg2020alt_appendix.htm")

```

## prepare over/underestimation plot
```{r accuracy per candidate}
# 2016
final_2016$accuracy_16t <- final_2016$rawpoll_trump - final_2016$trump_new
#summary(final_2016$accuracy_16t)

final_2016$accuracy_16c <- final_2016$rawpoll_clinton - final_2016$clinton_new
# summary(final_2016$accuracy_16c)

# 2020
final_2020c$accuracy_20t <- final_2020c$pct.x - final_2020c$trump_new
# summary(final_2016$accuracy_16t)

final_2020c$accuracy_20b <- final_2020c$pct.y - final_2020c$biden_new
# summary(final_2016$accuracy_16c)

```

```{r over/underestimation plot}
png(file="tables_figures_and_graphs/over_underestimation.png",
width=600, height=350)

par(mfrow = c(2, 2),
    mar = c(2, 0, 0, 0) + 2)
plot(
  density(final_2020c$accuracy_20t, na.rm=T),
  main =  "Trump 2020",
  xlim = c(-35, 15),
  xlab = "Signed margin of error")
plot(
  density(final_2020c$accuracy_20b, na.rm=T),
  main = "Biden 2020",
  xlim = c(-35, 15),
  xlab = "Signed margin of error")
plot(
  density(final_2016$accuracy_16t, na.rm=T),
  main =  "Trump 2016",
  xlim = c(-35, 15),
  xlab = "Signed margin of error")
plot(
  density(final_2016$accuracy_16c, na.rm=T),
  main = "Clinton 2016",
  xlim = c(-35, 15),
  xlab = "Signed margin of error")

dev.off()

```
